---
title: "R Notebook"
output: html_notebook
---

```{r}
rm(list = ls(all=T))
setwd("E:\\INSOFE RAR\\census")
train_data<- fread("census-income.data",na.strings = c(""," ","?","NA",NA))
test_data<-fread("census-income.test",na.strings = c(""," ","?","NA",NA))


sum(is.na(train_data))

```

```{r}
names<-c("age","classofworker","industrycode","occupationcode","education","wageperhour","enrolled.school","marital.status","major.industry","major.occupation","race","hispanic","sex","union","reason.unemployment","workstat","capgain","caploss","divstocks","taxfilerstatus","state.previous","region.previous","householdandfamily","household","instance.weight","MIGMTR1","MIGMTR3","MIGMTR4","livethishouse","sunbeltmigr","nbunder18","parent.present","birthfather","birthmother","birthself","citizenship","ownbusiness","questionnaireveteran","veteransbenefits","weeksworked","year","income")

names(train_data) <- names
names(test_data) <- names 
```

#Removing columns with null values percentage greater that 30

```{r}
train_data$MIGMTR1<-NULL
train_data$MIGMTR3<-NULL
train_data$MIGMTR4<-NULL
train_data$sunbeltmigr<-NULL


test_data$MIGMTR1<-NULL
test_data$MIGMTR3<-NULL
test_data$MIGMTR4<-NULL
test_data$sunbeltmigr<-NULL
```


```{r}
Numerical_trian=train_data[,c(1,3,4,6,17,18,19,25,27,33,35,36,37)]
categorical_train=train_data[,c(2,5,7:16,20:24,26,28:32,34,38)]


categorical_test=test_data[,c(2,5,7:16,20:24,26,28:32,34,38)]
Numerical_test=test_data[,c(1,3,4,6,17,18,19,25,27,33,35,36,37)]


colMeans(is.na(categorical_train))
sum(is.na(categorical_train))
categorical_train<-centralImputation(categorical_train)
categorical_test<-centralImputation(categorical_test)


sum(is.na(Numerical_trian))
colSums(is.na(Numerical_trian))
colSums(is.na(Numerical_test))

str(Numerical_trian)
str(categorical_train)
target<-train_data$income
# Numerical_trian$income<-as.numeric(target)
```

# In year 95, some variables are not present. Since we'll need to predict for both years in our test, we'll delete them.
```{r}
train_data$region.previous<-NULL
train_data$state.previous<-NULL
train_data$livethishouse<-NULL
train_data$year<-NULL
train_data$industrycode<-NULL
train_data$occupationcode<-NULL


test_data$region.previous<-NULL
test_data$state.previous<-NULL
test_data$livethishouse<-NULL
test_data$year<-NULL
test_data$industrycode<-NULL
test_data$occupationcode<-NULL

```

```{r}
categorical_train$income[categorical_train$income=='- 50000.']=0
categorical_train$income[categorical_train$income=='50000+.']=1


categorical_test$income[categorical_test$income=='- 50000.']=0
categorical_test$income[categorical_test$income=='50000+.']=1
```

```{r}

categorical_train$classofworker[categorical_train$classofworker== "Federal government"] =  "Governement"
categorical_train$classofworker[categorical_train$classofworker=="Local government"] = "Governement"
categorical_train$classofworker[categorical_train$classofworker=="State government"] = "Governement"
categorical_train$classofworker[categorical_train$classofworker=="Never worked"] = "Not-Working"
categorical_train$classofworker[categorical_train$classofworker=="Without pay"] =  "Not-Working"
categorical_train$classofworker[categorical_train$classofworker=="Self-employed-incorporated"] = "Self-Employed"
categorical_train$classofworker[categorical_train$classofworker=="Self-employed-not incorporated"] = "Self-Employed"
categorical_train$classofworker <- as.factor(categorical_train$classofworker)
summary(categorical_train$classofworker)



categorical_test$classofworker[categorical_test$classofworker== "Federal government"] =  "Governement"
categorical_test$classofworker[categorical_test$classofworker=="Local government"] = "Governement"
categorical_test$classofworker[categorical_test$classofworker=="State government"] = "Governement"
categorical_test$classofworker[categorical_test$classofworker=="Never worked"] = "Not-Working"
categorical_test$classofworker[categorical_test$classofworker=="Without pay"] =  "Not-Working"
categorical_test$classofworker[categorical_test$classofworker=="Self-employed-incorporated"] = "Self-Employed"
categorical_test$classofworker[categorical_test$classofworker=="Self-employed-not incorporated"] = "Self-Employed"
categorical_test$classofworker <- as.factor(categorical_test$classofworker)
summary(categorical_test$classofworker)
```

# We need to reduce the number of factors in order to have a convergent regression. 


```{r}


categorical_train$education<-as.character(categorical_train$education)

categorical_train$education[categorical_train$education == "10th grade"] = "No Degree"
categorical_train$education[categorical_train$education == "12th grade no diploma"] = "No Degree"
categorical_train$education[categorical_train$education == "5th or 6th grade"] = "No Degree"
categorical_train$education[categorical_train$education == "9th grade"] = "No Degree"
categorical_train$education[categorical_train$education=="Children"]="No Degree"
categorical_train$education[categorical_train$education=="High school graduate"]="No Degree"
categorical_train$education[categorical_train$education == "11th grade"] = "No Degree"
categorical_train$education[categorical_train$education == "1st 2nd 3rd or 4th grade"] = "No Degree"
categorical_train$education[categorical_train$education == "7th and 8th grade"] = "No Degree"
categorical_train$education[categorical_train$education == "Less than 1st grade"] = "No Degree"

categorical_train$education[categorical_train$education == "Some college but no degree"] = " High school graduate"

categorical_train$education[categorical_train$education== "Masters degree(MA MS MEng MEd MSW MBA)"]= "Masters"

categorical_train$education[categorical_train$education == "Associates degree-occup /vocational"] = "Associate"
categorical_train$education[categorical_train$education == "Associates degree-academic program"] = "Associate"
categorical_train$education[categorical_train$education == "Bachelors degree(BA AB BS)"] = "Graduate"
categorical_train$education[categorical_train$education == "Doctorate degree(PhD EdD)"] = "Doctorate"
categorical_train$education[categorical_train$education == "Prof school degree (MD DDS DVM LLB JD)"] = "Doctorate"

categorical_train$education<-as.factor(categorical_train$education)
summary(categorical_train$education)
levels(categorical_train$education)





categorical_test$education<-as.character(categorical_test$education)

categorical_test$education[categorical_test$education == "10th grade"] = "No Degree"
categorical_test$education[categorical_test$education == "12th grade no diploma"] = "No Degree"
categorical_test$education[categorical_test$education == "5th or 6th grade"] = "No Degree"
categorical_test$education[categorical_test$education == "9th grade"] = "No Degree"
categorical_test$education[categorical_test$education=="Children"]="No Degree"
categorical_test$education[categorical_test$education=="High school graduate"]="No Degree"
categorical_test$education[categorical_test$education == "11th grade"] = "No Degree"
categorical_test$education[categorical_test$education == "1st 2nd 3rd or 4th grade"] = "No Degree"
categorical_test$education[categorical_test$education == "7th and 8th grade"] = "No Degree"
categorical_test$education[categorical_test$education == "Less than 1st grade"] = "No Degree"

categorical_test$education[categorical_test$education == "Some college but no degree"] = " High school graduate"

categorical_test$education[categorical_test$education== "Masters degree(MA MS MEng MEd MSW MBA)"]= "Masters"

categorical_test$education[categorical_test$education == "Associates degree-occup /vocational"] = "Associate"
categorical_test$education[categorical_test$education == "Associates degree-academic program"] = "Associate"
categorical_test$education[categorical_test$education == "Bachelors degree(BA AB BS)"] = "Graduate"
categorical_test$education[categorical_test$education == "Doctorate degree(PhD EdD)"] = "Doctorate"
categorical_test$education[categorical_test$education == "Prof school degree (MD DDS DVM LLB JD)"] = "Doctorate"

categorical_test$education<-as.factor(categorical_test$education)
summary(categorical_test$education)
levels(categorical_test$education)

```

```{r}
categorical_train$citizenship<-as.character(categorical_train$citizenship)
categorical_train$citizenship[categorical_train$citizenship=="Native- Born in the United States"]="Native-Born"
categorical_train$citizenship[categorical_train$citizenship=="Foreign born- Not a citizen of U S"]="Foriegn_born"
categorical_train$citizenship[categorical_train$citizenship=="Native- Born abroad of American Parent(s)"]="Native-Born"
categorical_train$citizenship[categorical_train$citizenship=="Foreign born- U S citizen by naturalization"]="Foriegn_born"

categorical_train$citizenship[categorical_train$citizenship=="Native- Born in Puerto Rico or U S Outlying"]="Native-Born"
categorical_train$citizenship<-as.factor(categorical_train$citizenship)
levels(categorical_train$citizenship)


categorical_test$citizenship<-as.character(categorical_test$citizenship)
categorical_test$citizenship[categorical_test$citizenship=="Native- Born in the United States"]="Native-Born"
categorical_test$citizenship[categorical_test$citizenship=="Foreign born- Not a citizen of U S"]="Foriegn_born"
categorical_test$citizenship[categorical_test$citizenship=="Native- Born abroad of American Parent(s)"]="Native-Born"
categorical_test$citizenship[categorical_test$citizenship=="Foreign born- U S citizen by naturalization"]="Foriegn_born"

categorical_test$citizenship[categorical_test$citizenship=="Native- Born in Puerto Rico or U S Outlying"]="Native-Born"
categorical_test$citizenship<-as.factor(categorical_test$citizenship)
levels(categorical_test$citizenship)
```


<!-- ```{r} -->
<!-- str(categorical_train) -->
<!-- categorical$householdandfamily<-as.factor(categorical$householdandfamily) -->
<!-- categorical<-as.factor(categorical$questionnaireveteran) -->
<!-- categorical_train <- lapply(categorical_train , factor) -->
<!-- levels(categorical_train$marital.status) -->
<!-- levels(categorical_train$major.industry) -->


<!-- ``` -->
# In year 95, some variables are not present. Since we'll need to predict for both years in our test, we'll delete them.
```{r}
categorical_train$region.previous<-NULL
categorical_train$state.previous<-NULL
categorical_train$livethishouse<-NULL
categorical_train$year<-NULL
categorical_train$industrycode<-NULL
categorical_train$occupationcode<-NULL


categorical_test$region.previous<-NULL
categorical_test$state.previous<-NULL
categorical_test$livethishouse<-NULL
categorical_test$year<-NULL
categorical_test$industrycode<-NULL
categorical_test$occupationcode<-NULL


```


```{r}
categorical_train$marital.status <- as.character(categorical_train$marital.status)
categorical_train$marital.status[categorical_train$marital.status=="Divorced"] = "Separated"
categorical_train$marital.status[categorical_train$marital.status=="Married-spouse absent"] = "Separated"
categorical_train$marital.status <- as.factor(categorical_train$marital.status)
summary(categorical_train$marital.status)


categorical_test$marital.status <- as.character(categorical_test$marital.status)
categorical_test$marital.status[categorical_test$marital.status=="Divorced"] = "Separated"
categorical_test$marital.status[categorical_test$marital.status=="Married-spouse absent"] = "Separated"
categorical_test$marital.status <- as.factor(categorical_test$marital.status)
summary(categorical_test$marital.status)


```

```{r}
categorical_train$citizenship<-as.character(categorical_train$citizenship)
categorical_train$citizenship[categorical_train$citizenship=="Native- Born in the United States"]="Native-Born"
categorical_train$citizenship[categorical_train$citizenship=="Foreign born- Not a citizen of U S"]="Foriegn_born"
categorical_train$citizenship[categorical_train$citizenship=="Native- Born abroad of American Parent(s)"]="Native-Born"
categorical_train$citizenship[categorical_train$citizenship=="Foreign born- U S citizen by naturalization"]="Foriegn_born"

categorical_train$citizenship[categorical_train$citizenship=="Native- Born in Puerto Rico or U S Outlying"]="Native-Born"
categorical_train$citizenship<-as.factor(categorical_train$citizenship)
levels(categorical_train$citizenship)

categorical_test$citizenship<-as.character(categorical_test$citizenship)
categorical_test$citizenship[categorical_test$citizenship=="Native- Born in the United States"]="Native-Born"
categorical_test$citizenship[categorical_test$citizenship=="Foreign born- Not a citizen of U S"]="Foriegn_born"
categorical_test$citizenship[categorical_test$citizenship=="Native- Born abroad of American Parent(s)"]="Native-Born"
categorical_test$citizenship[categorical_test$citizenship=="Foreign born- U S citizen by naturalization"]="Foriegn_born"

categorical_test$citizenship[categorical_test$citizenship=="Native- Born in Puerto Rico or U S Outlying"]="Native-Born"
categorical_test$citizenship<-as.factor(categorical_test$citizenship)
levels(categorical_test$citizenship)

```
#Deleting "In group quarters which is not present in test data"

```{r}
categorical_train$householdandfamily<-as.character(categorical_train$householdandfamily)
categorical_train<-categorical_train[!(categorical_train$householdandfamily=="In group quarters"),]

categorical_test$household<-as.character(categorical_test$household)
categorical_test<-categorical_test[!(categorical_test$household=="Group Quarters- Secondary individual"),]


```




```{r}
str(categorical_train)
str(categorical_test)
categorical_train <- lapply(categorical_train , factor)
categorical_test <- lapply(categorical_test , factor)

categorical_train<-data.frame(categorical_train)
categorical_test<-data.frame(categorical_test)



 library(dummies)

categorical_train_d<-dummy.data.frame(categorical_train[,-22],sep =".")
categorical_train_d$income<-categorical_train$income
categorical_test_d<-dummy.data.frame(categorical_test[,-22],sep =".")
categorical_test_d$income<-categorical_test$income

Numerical_trian$income<-NULL
Std_model<-preProcess(Numerical_trian,method = c("center","scale"))

Numerical_trian <- predict(Std_model, Numerical_trian)
  
Numerical_test <- predict(Std_model, Numerical_test)

CN_train<-cbind(Numerical_trian,categorical_train_d)
CN_test<-cbind(Numerical_test,categorical_test_d)

levels(categorical_test$household)
levels(categorical_train$household)
```



```{r}
nearZ<-nearZeroVar(CN_train,saveMetrics = T)
nzv<-which(nearZ$nzv=="TRUE")
CN_train[,c(nzv)]<-NULL
CN_test[,c(nzv)]<-NULL

```







